Kindly Bent to Free Us

### 2019-11-15 kindly/lang.tex
Pattern matching.

\subsubsection{Environments and bindings}
\label{sdtyping:envs}
\begin{figure*}[tp]
  \begin{minipage}{0.28\linewidth}
    \begin{align*}
      \E &::=\ \Eempty \mid \E;B\tag{Environments}\\
      B &::=\ \bnone \tag{Empty} \\
         &\mid \bvar{x}{\schm}\tag{Variables}\\
         &\mid \bvar{\borrow{x}}{\borrowty{k}{\tau}}\tag{Borrows}\\
         &\mid \svar{x}{\schm}^n \tag{Suspended}\\
         &\mid \bvar{\tvar}{\kschm}\tag{Types}
         % |&\ \tydecl{t}{\kschm}{K}{\tau}\tag{Type Declaration}\\
    \end{align*}
    \caption{Type environments}
    \label{grammar:env}
  \end{minipage}\hfill
  \begin{minipage}{0.7\linewidth}
    \begin{mathpar}
      \ruleSDBorrow
      \and
      \ruleSDLam
      \and
      \ruleSDVar
      \and
      \ruleSDPair
      \and
      \ruleSDRegion
      \and
      \ruleSDMatchPair
    \end{mathpar}
    \caption{Selected typing rules}
    \label{selectrules:borrow}
    \label{selectrules:binders}
    \label{sdtyp:pair}
    \label{selectrules:region}
  \end{minipage}
\end{figure*}

@@

@@

\subsubsection{Pattern matching}

\TODO{Present this if we have enough space}

>>>>

\subsubsection{Environments and bindings}
\label{sdtyping:envs}
\begin{figure*}[tp]
  \begin{minipage}{0.28\linewidth}
    \begin{align*}
      \E &::=\ \Eempty \mid \E;B\tag{Environments}\\
      B &::=\ \bnone \tag{Empty} \\
         &\mid \bvar{x}{\schm}\tag{Variables}\\
         &\mid \bvar{\borrow{x}}{\borrowty{k}{\tau}}\tag{Borrows}\\
         &\mid \svar{x}{\schm}^n \tag{Suspended}\\
         &\mid \bvar{\tvar}{\kschm}\tag{Types}
         % |&\ \tydecl{t}{\kschm}{K}{\tau}\tag{Type Declaration}\\
    \end{align*}
    \caption{Type environments}
    \label{grammar:env}
  \end{minipage}\hfill
  \begin{minipage}{0.7\linewidth}
    \begin{mathpar}
      \ruleSDBorrow
      \and
      \ruleSDLam
      \and
      \ruleSDVar
      \and
      \ruleSDPair
      \and
      \ruleSDRegion
      \and
      \ruleSDMatchPair
    \end{mathpar}
    \caption{Selected typing rules}
    \label{selectrules:borrow}
    \label{selectrules:binders}
    \label{sdtyp:pair}
    \label{selectrules:region}
    \label{selectrules:matching}
  \end{minipage}
\end{figure*}

@@

\subsubsection{Pattern matching}
\label{sdtyping:matching}

In order to deconstruct pairs in the presence of linear types, we require
more than projection operators. Indeed, given a linear pair $x$,
$\operatorname{fst} x$ would consume $x$ without giving access to the second
member.
Furthermore, we would like to be able to deconstruct a borrow of a pair
and obtain borrows of the components.
For these purposes, we provide the pattern matching construct
$\matchin{x,x'}{e_1}{e_2}$. The typing rule is shown in \cref{selectrules:matching}.
A syntactic marker $\etransfm$ indicates if it applies to
a pair ($\etransfm = \operatorname{id}$) or a borrow ($\etransfm = \&^\BORROW$).
If $\etransfm = \operatorname{id}$, it operators a regular destruct on pair.
Otherwise, $e_1$ is expected to be of type
$\borrowty{k}{\tyPair{\tau_1}{\tau'_1}}$
while $x$ and $x'$ are of type respectively
$\borrowty{k}{\tau_1}$ and $\borrowty{k}{\tau'_1}$.

The typing of pattern matching is voluntarily simple here.
Using a typeclass mechanism could avoid the need for explicit markers
for borrow matches. Furthermore, a richer pattern language would allow
finer-grained borrowing of inner components.
We believe both extension are compatible with our system
and are out of scope of this article.

@@
